Descriptive statistics

### 2023-10-06 Code/0_PrepareTrainingData.Rmd
Final edits

# load data file
original.train.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/training/Dataset Predictiemodel.sav",
                                    to.data.frame=TRUE, origin="01-01-1970")

@@

### Inclusion criteria
Exclude when Exclusion = 2: based on duplicates, urogenital melanoma, etc. Note:
Leave the patients included that have EXCLUSIE = 1 and Dbstatus = 8 (Dbstatus=0: 
negative, Dbstatus=1: positive), even though diagnosis and operation date are 
very close.
```{r, eval=TRUE}
sel.train.data <- original.train.data[original.train.data$EXCLUSIE!="Exclusion", selected.vars]
cat("Number of patients excluded:", nrow(original.train.data)-nrow(sel.train.data), "\n")
```
Omit one patient that has no last follow-up date
```{r, eval=TRUE, echo=FALSE}
sel.train.data <- sel.train.data[!is.na(sel.train.data$Last_FU),]
```

@@

#### Describe years
```{r, eval=TRUE}
dates <- as.POSIXct(sel.train.data$SNdate, origin="1582-10-14")
Years <- as.numeric(format(dates, format="%Y"))
utils::write.table(cbind(table(Years), table(Years, sel.train.data$Recurrence)[,2], 
                         table(Years, sel.train.data$MSM)[, 2]),
                   file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Time.Rec.MSM.txt"),
                   row.names=TRUE, sep=",")
hist(Years, main="")
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM")
training.data.to.be.imputed <- sel.train.data[, variables.to.be.imputed]
```

@@

Relation between outcome and covariates
```{r, eval=TRUE}
training.data.to.be.imputed$logRec.time <- log(sel.train.data$Rec.time+1/365.25)
training.data.to.be.imputed$logFU.time <- log(sel.train.data$FU.time+1/365.25)
missings.table(training.data.to.be.imputed)
save(list=c("dd.training", "training.data.to.be.imputed"),file="Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata", compress=TRUE)
```

>>>>

# load data file
original.training.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/training/Dataset Predictiemodel.sav",
                                    to.data.frame=TRUE, origin="01-01-1970")

@@

### Inclusion criteria
Exclude when Exclusion = 2: based on duplicates, urogenital melanoma, etc. Note:
Leave the patients included that have EXCLUSIE = 1 and Dbstatus = 8 (Dbstatus=0: 
negative, Dbstatus=1: positive), even though diagnosis and operation date are 
very close.
```{r, eval=TRUE}
sel.train.data <- original.training.data[original.training.data$EXCLUSIE!="Exclusion", selected.vars]
cat("Number of patients excluded:", nrow(original.training.data)-nrow(sel.train.data), "\n")
```
Omit one patient that has no last follow-up date
```{r, eval=TRUE, echo=FALSE}
sel.train.data <- sel.train.data[!is.na(sel.train.data$Last_FU),]
```

@@

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis", "Status", "Recurrence", "dead", "MSM")
training.data.to.be.imputed <- sel.train.data[, variables.to.be.imputed]
```

@@

Relation between outcome and covariates
```{r, eval=TRUE}
training.data.to.be.imputed$logRec.time <- log(sel.train.data$Rec.time+1/365.25)
training.data.to.be.imputed$logFU.time <- log(sel.train.data$FU.time+1/365.25)
missings.table(training.data.to.be.imputed)
# save(list=c("dd.training", "training.data.to.be.imputed"),file="Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata", compress=TRUE)
```

### 2023-10-06 Code/0_PrepareValidationData.Rmd
Final edits

#### Melanoma specific mortality (MSM)
```{r, eval=TRUE}
sel.val.data$dead <- as.numeric(stringr::str_detect(sel.val.data$Status, "Dead"))
sel.val.data$dead <- factor(sel.val.data$dead, labels=c("No", "Yes"))
table(sel.val.data$Status, sel.val.data$dead)
sel.val.data$MSM <- as.numeric(sel.val.data$Status=="Dead, melanoma       ")
```
#### Mitosis
Define Mitosis as Yes/No
```{r, eval=TRUE}
sel.val.data$Mitosis <- as.numeric(!is.na(sel.val.data$Mitosis))
```

@@

#### Set levels
```{r, eval=TRUE}
sel.val.data$Ulceration <- factor(sel.val.data$Ulceration, labels=c("No", "Yes"))
sel.val.data$Mitosis <- factor(sel.val.data$Mitosis, labels=c("Yes", "No"))

>>>>

#### Melanoma specific mortality (MSM)
```{r, eval=TRUE}
sel.val.data$dead <- as.numeric(stringr::str_detect(sel.val.data$Status, "Dead"))
sel.val.data$dead <- factor(sel.val.data$dead, labels=c("No", "Yes"))
table(sel.val.data$Status, sel.val.data$dead)
sel.val.data$MSM <- as.numeric(sel.val.data$Status=="Dead, melanoma       ")
```
#### Mitosis
Define Mitosis as Yes/No
```{r, eval=TRUE}
sel.val.data$Mitosis <- as.numeric(sel.val.data$Mitosis)
```

@@

#### Set levels
```{r, eval=TRUE}
sel.val.data$Ulceration <- factor(sel.val.data$Ulceration, labels=c("No", "Yes"))

### 2023-10-06 Code/10_AJCC.Rmd
Final edits

# load libraries
library(PredictionTools)  # to plot calibration
library(compareC)         # compare AJCC to model
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

>>>>

# load libraries
library(PredictionTools)  # to plot calibration
library(compareC)         # compare AJCC to model
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      if (type=="training"){
        mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      } else{
        mitosis.patient <- ifelse(imputation.i[patient, "Mitosis"]<1, "Yes", "No")
      }
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

### 2023-10-06 Code/1_DescriptivesTable.Rmd
Update
Final edits

load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata")
load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata")
descr.table <- arsenal::tableby(Center ~ SNstatus
                                +Sex
                                +Age.SN
                                +arsenal::includeNA(Ulceration)
                                +arsenal::includeNA(Loc_CAT)
                                +arsenal::includeNA(Histology)
                                +Breslow
                                +multiple.fields
                                +Tot_SNs_neg
                                +Tot_SNs_pos
                                +Rdamcrit
                                +arsenal::includeNA(Dewar)
                                +arsenal::includeNA(Mitosis)
                                +Recurrence
                                +dead
                                +MSM
                                +logRec.time
                                +logFU.time, data=rbind(training.data.to.be.imputed, validation.data.to.be.imputed), # training.data.to.be.imputed,
                                total.pos="before",
                                digits = 1,
                                test = F,
                                total = T,
                                numeric.test = "kwt", cat.test = "chisq",
                                numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss2"),
                                cat.stats = c("countpct"),
                                stats.labels = list(
                                  meansd = "Mean (SD)",
                                  medianq1q3 = "Median (Q1, Q3)",
                                  range = "Min - Max",
                                  Nmiss2 = "Missing"
                                ))
my_labels <- list(
  SNstatus = "SN status",
  Age.SN = "Age",
  Loc_CAT = "Location",
  multiple.fields = "Multiple fields",
  Tot_SNs_neg = "Total number of negative SN",
  Tot_SNs_pos = "Total number of positive SN",
  Rdamcrit = "SN tumour burden",
  Dewar = "Location metastasis in lymph",
  dead = "Deaths",
  MSM = "Melanoma specific mortality"
)
```

Table 1. Descriptive statistics.
```{r echo=FALSE}
knitr::kable(summary(descr.table, labelTranslations=my_labels))
```

>>>>

load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata")
load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata")
validation.data.to.be.imputed$Mitosis <- validation.data.to.be.imputed$Mitosis>0
validation.data.to.be.imputed$Mitosis <- factor(validation.data.to.be.imputed$Mitosis, labels=c("No", "Yes"))
pooled.data.to.be.imputed <- rbind(training.data.to.be.imputed, validation.data.to.be.imputed)
pooled.data.to.be.imputed$Continent <- c(rep("Europe", nrow(training.data.to.be.imputed)), 
                                         rep("Australie", nrow(validation.data.to.be.imputed)))

descr.table <- arsenal::tableby(Continent ~ SNstatus
                                +Sex
                                +Age.SN
                                +arsenal::includeNA(Ulceration)
                                +arsenal::includeNA(Loc_CAT)
                                +arsenal::includeNA(Histology)
                                +Breslow
                                +multiple.fields
                                +arsenal::includeNA(Mitosis)
                                +Recurrence
                                +dead
                                +MSM
                                +logRec.time
                                +logFU.time, 
                                data=pooled.data.to.be.imputed,
                                total.pos="before",
                                digits = 1,
                                test = F,
                                total = F,
                                numeric.test = "kwt", cat.test = "chisq",
                                numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss2"),
                                cat.stats = c("countpct"),
                                stats.labels = list(
                                  meansd = "Mean (SD)",
                                  medianq1q3 = "Median (Q1, Q3)",
                                  range = "Min - Max",
                                  Nmiss2 = "Missing"
                                ))

descr.table.pos <- arsenal::tableby(Continent ~ Tot_SNs_pos
                                +Rdamcrit
                                +arsenal::includeNA(Dewar),
                                data=pooled.data.to.be.imputed[pooled.data.to.be.imputed$SNstatus=="Positive",],
                                total.pos="before",
                                digits = 1,
                                test = F,
                                total = T,
                                numeric.test = "kwt", cat.test = "chisq",
                                numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss2"),
                                cat.stats = c("countpct"),
                                stats.labels = list(
                                  meansd = "Mean (SD)",
                                  medianq1q3 = "Median (Q1, Q3)",
                                  range = "Min - Max",
                                  Nmiss2 = "Missing"
                                ))

descr.table.neg <- arsenal::tableby(Continent ~ Tot_SNs_neg,
                                data=pooled.data.to.be.imputed[pooled.data.to.be.imputed$SNstatus=="Negative",],
                                total.pos="before",
                                digits = 1,
                                test = F,
                                total = T,
                                numeric.test = "kwt", cat.test = "chisq",
                                numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss2"),
                                cat.stats = c("countpct"),
                                stats.labels = list(
                                  meansd = "Mean (SD)",
                                  medianq1q3 = "Median (Q1, Q3)",
                                  range = "Min - Max",
                                  Nmiss2 = "Missing"
                                ))
my_labels <- list(
  SNstatus = "SN status",
  Age.SN = "Age",
  Loc_CAT = "Location",
  multiple.fields = "Multiple fields",
  Tot_SNs_neg = "Total number of negative SN",
  Tot_SNs_pos = "Total number of positive SN",
  Rdamcrit = "SN tumour burden",
  Dewar = "Location metastasis in lymph",
  dead = "Deaths",
  MSM = "Melanoma specific mortality"
)
```

Table 1. Descriptive statistics.
```{r echo=FALSE}
knitr::kable(summary(descr.table, labelTranslations=my_labels))

# only negative
knitr::kable(summary(descr.table.neg, labelTranslations=my_labels))

# only positive
knitr::kable(summary(descr.table.pos, labelTranslations=my_labels))
```

### 2023-10-06 Code/3_Descriptives.Rmd
Final edits

  save(list=c(paste0("dd.", type), "m", 
              paste0(type, ".data.to.be.imputed"), 
              paste0("imputed.", type, ".data"), 
              paste0("last.imputed.", type, ".data.set"), 
              paste0("S.Rec.", type), 
              paste0("S.MSM.", type)),
       file=paste0("Z:/Project Melanoom/PaperMelanoma/Data/", type, "/", type, ".data.Rdata"), compress=TRUE)
}
prepare.data(type="training")
prepare.data(type="validation")
```

### Kaplan-Meier plots
```{r, eval=TRUE}
load(paste0("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData"))
survminer::ggsurvplot(survival::survfit(S.Rec.training ~ Center, data=last.imputed.training.data.set), conf.int=TRUE, risk.table=TRUE)
survminer::ggsurvplot(survival::survfit(S.MSM.training ~ Center, data=last.imputed.training.data.set), conf.int=TRUE, risk.table=TRUE)
load(paste0("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData"))
survminer::ggsurvplot(survival::survfit(S.Rec.validation ~ 1, data=last.imputed.validation.data.set), conf.int=TRUE, risk.table=TRUE)
survminer::ggsurvplot(survival::survfit(S.MSM.validation ~ 1, data=last.imputed.validation.data.set), conf.int=TRUE, risk.table=TRUE)
```

# Median time to recurrence and follow-up time
```{r, eval=TRUE}
median.time <- function(data=NA, type=NA){
  # time in total
  if (type=="Rec"){
    median.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logRec.time), Recurrence=="recurred")~1, data=data[data$Center!="Australia",], reverse=TRUE))
    median.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logRec.time), Recurrence=="recurred")~Center, data=data, reverse=TRUE))
  } else if (type=="FU"){
    median.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM=="Death by melanoma")~1, data=data[data$Center!="Australia",], reverse=TRUE))
    median.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM=="Death by melanoma")~Center, data=data, reverse=TRUE))
  }

  # save in tab.time
  tab.time <- c("Total Development", paste(sprintf("%.2f", median.time.total$quantiles.survival[3, "quantile"]),
                   paste0("[", sprintf("%.2f", median.time.total$quantiles.survival[4, "quantile"]), "; ",
                          sprintf("%.2f", median.time.total$quantiles.survival[2, "quantile"]), "]")))

  # time per center
  for (center in c("Amsterdam", "Berlin", "Rotterdam", "Warsaw", "Australia")){
    med.table <- eval(parse(text=paste0("median.time$`Center=", center, "`")))
    med <- med.table[3,"quantile"]
    lower <- med.table[4,"quantile"]
    upper <- med.table[2,"quantile"]
    tab.time <- rbind(tab.time, c(center, paste(sprintf("%.2f", med), paste0("[", sprintf("%.2f", lower), "; ", sprintf("%.2f", upper), "]"))))
  }

  return(tab.time)
}
med.Rec <- median.time(data=rbind(training.data.to.be.imputed[, c("logRec.time", "Recurrence", "Center")], 
                       validation.data.to.be.imputed[, c("logRec.time", "Recurrence", "Center")]), type="Rec")
med.FU <- median.time(data=rbind(training.data.to.be.imputed[, c("logFU.time", "MSM", "Center")], 
                       validation.data.to.be.imputed[, c("logFU.time", "MSM", "Center")]), type="FU")
# save to .txt
write.table(cbind(c("Recurrence time", "Follow-up time"), 
                  rbind(c(med.Rec[, -1]), c(med.FU[,-1]))),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/med.table.txt"),
              row.names=FALSE, sep=",")
```

# Kaplan-Meier estimates of recurrence and MSM at 5 years
```{r, eval=TRUE}
Rec.MSM.table <- function(horizon=5, type=NULL, data.to.be.imputed=NULL){
  # all patients, Recurrence
  S.Rec <- eval(parse(text=paste0("S.Rec.", type)))
  SF.Rec.all <- survival::survfit(S.Rec~1, data=data.to.be.imputed)
  Rec.all <- 100*(1-SF.Rec.all$surv[SF.Rec.all$time==max(SF.Rec.all$time[SF.Rec.all$time<=horizon])])
  n.Rec.all <- sum(SF.Rec.all$n.event[SF.Rec.all$time<=horizon])

  # only negative patients, Recurrence
  SF.Rec.neg <-survival::survfit(S.Rec[data.to.be.imputed$SNstatus=="Negative"]~1, data=data.to.be.imputed)
  Rec.neg <- 100*(1-SF.Rec.neg$surv[SF.Rec.neg$time==max(SF.Rec.neg$time[SF.Rec.neg$time<=horizon])])
  n.Rec.neg <- sum(SF.Rec.neg$n.event[SF.Rec.neg$time<=horizon])

  # only positive patients, Recurrence
  SF.Rec.pos <- survival::survfit(S.Rec[data.to.be.imputed$SNstatus=="Positive"]~1, data=data.to.be.imputed)
  Rec.pos <- 100*(1-SF.Rec.pos$surv[SF.Rec.pos$time==max(SF.Rec.pos$time[SF.Rec.pos$time<=horizon])])
  n.Rec.pos <- sum(SF.Rec.pos$n.event[SF.Rec.pos$time<=horizon])

  # all patients, MSM
  S.MSM <- eval(parse(text=paste0("S.MSM.", type)))
  SF.MSM.all <- survival::survfit(S.MSM~1, data=data.to.be.imputed)
  MSM.all <- 100*(1-SF.MSM.all$surv[SF.MSM.all$time==max(SF.MSM.all$time[SF.MSM.all$time<=horizon])])
  n.MSM.all <- sum(SF.MSM.all$n.event[SF.MSM.all$time<=horizon])

  # negative patients, MSM
  SF.MSM.neg <- survival::survfit(S.MSM[data.to.be.imputed$SNstatus=="Negative"]~1, data=data.to.be.imputed)
  MSM.neg <- 100*(1-SF.MSM.neg$surv[SF.MSM.neg$time==max(SF.MSM.neg$time[SF.MSM.neg$time<=horizon])])
  n.MSM.neg <- sum(SF.MSM.neg$n.event[SF.MSM.neg$time<=horizon])

  # positive patients, MSM
  SF.MSM.pos <- survival::survfit(S.MSM[data.to.be.imputed$SNstatus=="Positive"]~1, data=data.to.be.imputed)
  MSM.pos <- 100*(1-SF.MSM.pos$surv[SF.MSM.pos$time==max(SF.MSM.pos$time[SF.MSM.pos$time<=horizon])])
  n.MSM.pos <- sum(SF.MSM.pos$n.event[SF.MSM.pos$time<=horizon])

  # merge in one table
  Rec.MSM.table <- data.frame(n=c(nrow(data.to.be.imputed),
                                  sum(data.to.be.imputed$SNstatus=="Negative"),
                                  sum(data.to.be.imputed$SNstatus=="Positive")),
                              e1=rep("", 3),
                              n.Rec=c(n.Rec.all, n.Rec.pos, n.Rec.neg),
                              n.MSM=c(n.MSM.all, n.MSM.pos, n.MSM.neg),
                              e2=rep("", 3),
                              Rec=sprintf("%.1f", c(Rec.all, Rec.neg, Rec.pos)),
                              MSM=sprintf("%.1f", c(MSM.all, MSM.neg, MSM.pos)))
  row.names(Rec.MSM.table) <- c("All", "Negative", "Positive")

@@

    assign(paste0("S.", outcome, ".5.", type), S.5)
  }
  save(list=c(paste0("dd.", type), "m", "horizon", 
              paste0(type, ".data.to.be.imputed"),
              paste0("imputed.", type, ".data"), 
              paste0("last.imputed.", type, ".data.set"),
              paste0("S.Rec.", type), 
              paste0("S.MSM.", type),
              paste0("S.Rec.5.", type), 
              paste0("S.MSM.5.", type)),
     file=paste0("Z:/Project Melanoom/PaperMelanoma/Data/", type, "/", type, ".data.Rdata"), compress=TRUE)
}
```

>>>>

  save(list=c(paste0("dd.", type), "m", 
              paste0(type, ".data.to.be.imputed"), 
              paste0("imputed.", type, ".data"), 
              paste0("last.imputed.", type, ".data.set"), 
              paste0("S.Rec.", type), 
              paste0("S.MSM.", type)),
       file=paste0("Z:/Project Melanoom/PaperMelanoma/Data/", type, "/", type, ".data.Rdata"), compress=TRUE)
}
prepare.data(type="training")
prepare.data(type="validation")
```
# Median time to recurrence and follow-up time
```{r, eval=TRUE}
median.time <- function(data=NA, type=NA){
  # time in total
  if (type=="Rec"){
    median.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logRec.time), Recurrence=="recurred")~1, data=data[data$Center!="Australia",], reverse=TRUE))
    median.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logRec.time), Recurrence=="recurred")~Center, data=data, reverse=TRUE))
  } else if (type=="FU"){
    median.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM=="Death by melanoma")~1, data=data[data$Center!="Australia",], reverse=TRUE))
    median.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM=="Death by melanoma")~Center, data=data, reverse=TRUE))
  }
  # save in tab.time
  tab.time <- c("Total Development", paste(sprintf("%.2f", median.time.total$quantile[3]),
                   paste0("[", sprintf("%.2f", median.time.total$quantile[4]), "; ",
                          sprintf("%.2f", median.time.total$quantile[2]), "]")))

  # time per center
  for (center in c("Amsterdam", "Berlin", "Rotterdam", "Warsaw", "Australia")){
    sel.center <- eval(parse(text=paste0("median.time$Center==\"", center, "\"")))
    q.center <- median.time$quantile[sel.center]
    med <- q.center[3]
    lower <- q.center[4]
    upper <- q.center[2]
    tab.time <- rbind(tab.time, c(center, paste(sprintf("%.2f", med), paste0("[", sprintf("%.2f", lower), "; ", sprintf("%.2f", upper), "]"))))
  }

  return(tab.time)
}
load(paste0("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData"))
load(paste0("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData"))
med.Rec <- median.time(data=rbind(training.data.to.be.imputed[, c("logRec.time", "Recurrence", "Center")], 
                       validation.data.to.be.imputed[, c("logRec.time", "Recurrence", "Center")]), type="Rec")
med.FU <- median.time(data=rbind(training.data.to.be.imputed[, c("logFU.time", "MSM", "Center")], 
                       validation.data.to.be.imputed[, c("logFU.time", "MSM", "Center")]), type="FU")
# save to .txt
write.table(cbind(c("Recurrence time", "Follow-up time"), 
                  rbind(c(med.Rec[, -1]), c(med.FU[,-1]))),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/med.table.txt"),
              row.names=FALSE, sep=",")
```

# Kaplan-Meier estimates of recurrence and MSM at 5 years
```{r, eval=TRUE}
Rec.MSM.table <- function(horizon=5, type=NULL, data.to.be.imputed=NULL){
  # all patients, Recurrence
  S.Rec <- eval(parse(text=paste0("S.Rec.", type)))
  SF.Rec.all <- survival::survfit(S.Rec~1, data=data.to.be.imputed)
  RFS.all <- 100*SF.Rec.all$surv[SF.Rec.all$time==max(SF.Rec.all$time[SF.Rec.all$time<=horizon])]
  n.Rec.all <- sum(SF.Rec.all$n.event[SF.Rec.all$time<=horizon])

  # only negative patients, Recurrence
  SF.Rec.neg <-survival::survfit(S.Rec[data.to.be.imputed$SNstatus=="Negative"]~1, data=data.to.be.imputed)
  RFS.neg <- 100*SF.Rec.neg$surv[SF.Rec.neg$time==max(SF.Rec.neg$time[SF.Rec.neg$time<=horizon])]
  n.Rec.neg <- sum(SF.Rec.neg$n.event[SF.Rec.neg$time<=horizon])

  # only positive patients, Recurrence
  SF.Rec.pos <- survival::survfit(S.Rec[data.to.be.imputed$SNstatus=="Positive"]~1, data=data.to.be.imputed)
  RFS.pos <- 100*SF.Rec.pos$surv[SF.Rec.pos$time==max(SF.Rec.pos$time[SF.Rec.pos$time<=horizon])]
  n.Rec.pos <- sum(SF.Rec.pos$n.event[SF.Rec.pos$time<=horizon])

  # all patients, MSM
  S.MSM <- eval(parse(text=paste0("S.MSM.", type)))
  SF.MSM.all <- survival::survfit(S.MSM~1, data=data.to.be.imputed)
  MSS.all <- 100*SF.MSM.all$surv[SF.MSM.all$time==max(SF.MSM.all$time[SF.MSM.all$time<=horizon])]
  n.MSM.all <- sum(SF.MSM.all$n.event[SF.MSM.all$time<=horizon])

  # negative patients, MSM
  SF.MSM.neg <- survival::survfit(S.MSM[data.to.be.imputed$SNstatus=="Negative"]~1, data=data.to.be.imputed)
  MSS.neg <- 100*SF.MSM.neg$surv[SF.MSM.neg$time==max(SF.MSM.neg$time[SF.MSM.neg$time<=horizon])]
  n.MSM.neg <- sum(SF.MSM.neg$n.event[SF.MSM.neg$time<=horizon])

  # positive patients, MSM
  SF.MSM.pos <- survival::survfit(S.MSM[data.to.be.imputed$SNstatus=="Positive"]~1, data=data.to.be.imputed)
  MSS.pos <- 100*SF.MSM.pos$surv[SF.MSM.pos$time==max(SF.MSM.pos$time[SF.MSM.pos$time<=horizon])]
  n.MSM.pos <- sum(SF.MSM.pos$n.event[SF.MSM.pos$time<=horizon])

  # merge in one table
  Rec.MSM.table <- data.frame(n=c(nrow(data.to.be.imputed),
                                  sum(data.to.be.imputed$SNstatus=="Negative"),
                                  sum(data.to.be.imputed$SNstatus=="Positive")),
                              e1=rep("", 3),
                              n.Rec=c(n.Rec.all, n.Rec.pos, n.Rec.neg),
                              n.MSM=c(n.MSM.all, n.MSM.pos, n.MSM.neg),
                              e2=rep("", 3),
                              RFS=sprintf("%.1f", c(RFS.all, RFS.neg, RFS.pos)),
                              MSS=sprintf("%.1f", c(MSS.all, MSS.neg, MSS.pos)))
  row.names(Rec.MSM.table) <- c("All", "Negative", "Positive")

@@

    assign(paste0("S.", outcome, ".5.", type), S.5)
  }
  save(list=c(paste0("dd.", type), "m", "horizon", 
              paste0(type, ".data.to.be.imputed"),
              paste0("imputed.", type, ".data"), 
              paste0("last.imputed.", type, ".data.set"),
              paste0("S.Rec.", type), 
              paste0("S.MSM.", type),
              paste0("S.Rec.5.", type), 
              paste0("S.MSM.5.", type)),
     file=paste0("Z:/Project Melanoom/PaperMelanoma/Data/", type, "/", type, ".data.Rdata"), compress=TRUE)
}
```
### Kaplan-Meier plots
```{r, eval=TRUE}
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/KM.Rec.training.png"), width=16, height=16, units="cm", res=300)
survminer::ggsurvplot(survival::survfit(S.Rec.5.training ~ Center, data=last.imputed.training.data.set), 
                      legend.title="",
                      legend.labs=c("Rotterdam", "Berlin", "Warsaw", "Amsterdam"),
                      title="Development",
                      ylab="Recurrence-free survival probability",
                      xlab="Time in years",
                      conf.int=TRUE, risk.table=TRUE)+
  ggplot2::guides(colour=guide_legend(nrow=2))
grDevices::dev.off()
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/KM.MSM.training.png"), width=16, height=16, units="cm", res=300)
survminer::ggsurvplot(survival::survfit(S.MSM.5.training ~ Center, data=last.imputed.training.data.set), 
                      legend.title="",
                      legend.labs=c("Rotterdam", "Berlin", "Warsaw", "Amsterdam"),
                      title="Development",
                      ylab="Melanoma-specific survival probability",
                      xlab="Time in years",
                      conf.int=TRUE, risk.table=TRUE)+
  ggplot2::guides(colour=guide_legend(nrow=2))
grDevices::dev.off()

load(paste0("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData"))
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/KM.Rec.validation.png"), width=16, height=16, units="cm", res=300)
survminer::ggsurvplot(survival::survfit(S.Rec.5.validation ~ 1, data=last.imputed.validation.data.set), 
                      legend.title="",
                      legend.labs="Australia",
                      title="Validation",
                      ylab="Recurrence-free survival probability",
                      xlab="Time in years",
                      conf.int=TRUE, risk.table=TRUE)
grDevices::dev.off()
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/KM.MSM.validation.png"), width=16, height=16, units="cm", res=300)
survminer::ggsurvplot(survival::survfit(S.MSM.5.validation ~ 1, data=last.imputed.validation.data.set), 
                      legend.title="",
                      legend.labs="Australia",
                      title="Validation",
                      ylab="Melanoma-specific survival probability",
                      xlab="Time in years",
                      conf.int=TRUE, risk.table=TRUE)
grDevices::dev.off()
```
